---
title: "GEVP energy level extraction"
author: "Carsten Urbach"
output:
  pdf_document:
    citation_package: biblatex
    toc: yes
    
bibliography: gevp.bib
link_citation: yes
vignette: >
  %\VignetteIndexEntry{GEVP energy level extraction}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r echo=FALSE}
library(knitr)
library(hadron)
```

# Data Type \texttt{cf} for Correlation Functions

\texttt{hadron} provides a data type or better class for correlation
functions and correlation matrices, which is called \texttt{cf}. There
is a whole list of input
routines available to import data from HDF5, text or binary formats
into the \texttt{cf} container. The most important ones are compiled
in the following table:

```{r, echo=FALSE}
knitr::kable(data.frame(c("readtextcf", "readbinarycf", "readbinarysamples", "more?"),
                        c("text", "binary, HDF5", "binary", "...")),
             col.names=c("Hadron Function", "Correlator Format"))
```

For even more flexibility there is the \texttt{raw\_cf} container, for
which we refer to the documentation. 

In oder to solve the generalised eigenvalue problem (GEVP) one has to
read several correlation functions into one \texttt{cf} correlator
matrix. For this purpose the combine operation \texttt{c} is defined
for the class \texttt{cf}. Thus, for instance the following code
snipped can be used:

```{r, eval=FALSE}
Time <- 48
correlatormatrix <- cf()
for(i in c(1:4)) {
  tmp <- readbinarycf(files=paste0("corr", i, ".dat"), T=Time)
  correlatormatrix <- c(correlatormatrix, tmp)
}
rm(tmp)
```

This code snippet reads a correlator matrix with four correlation
functions from four files. The read functions can also directly read
from a list of files. File lists ca be created conveniently using the
following routines 

```{r, eval=FALSE}
getorderedfilelist <- function(path="./", basename="onlinemeas",
                               last.digits=4, ending="")
getconfignumbers <- function(ofiles, basename="onlinemeas",
                             last.digits=4, ending="")
getorderedconfigindices <- function(path="./", basename="onlinemeas",
                                    last.digits=4, ending="")
```

for which we refer to the documentation. 

# Resampling Strategies

Once the bare data is available as a \texttt{cf}, one has to decide
for an error analysis strategy. This can be either the bootstrap or
the jackknife. To demonstrate this we first load the sample
correlation matrix provided by hadron

```{r}
data(correlatormatrix)
```

which corresponds to a $2\times 2$ local-fuzzed correlator matrix with
quantum numbers of the pion.
First the resampling needs to be performed, for instance for the
(blocked) bootstrap

```{r, cache=TRUE}
boot.R <- 1500
boot.l <- 1
seed <- 1433567
correlatormatrix <- bootstrap.cf(cf=correlatormatrix,
                                 boot.R=boot.R,
                                 boot.l=boot.l,
                                 seed=seed)
```

Analogously, \texttt{jackknife.cf} initiates the jackknife
resampling. \texttt{boot.R} is the number of bootstrap replicates,
\texttt{boot.l} the block lentgh. Now, it is also possible to plot the
data with errors 

```{r, warning=FALSE}
plot(correlatormatrix, log="y",
     xlab=c("t/a"), ylab="C(t)")
```

# Solving the GEVP

Let us denote the correlator matrix by $C(t)$. Now we are going to
solve the generalised eigenvalue problem
\[
C(t)\, v_i(t, t_0)\ =\ \lambda_i(t, t_0)\, C(t_0)\, v_i(t, t_0)
\]
with some reference time value $t_0$. One can show that the so-called
principal correlators $\lambda(t, t_0)$ follow for large $t$-values
the following behaviour
\[
\lambda_i(t, t_0)\ \propto\ e^{-E_i(t-t_0)} + e^{-E_i(T-t+t_0)}\,.
\]
Here, $T$ is the time extend and we focus on a symmetric correlation
matrix in time. However, analogously one can show this with a minus
sign for anti-symmetric correlation matrices in time. Of course, we
also have $\lambda(t_0, t_0) = 1$. We re-write the generalised
eigenvalue problem by defining
\[
w_i\ =\ \sqrt{C(t_0)} v_i
\]
and solve the simple eigenvalue problem
\[
\sqrt{C(t_0)}^{-1}\,C(t)\,\sqrt{C(t_0)}^{-1}\, w_i\ =\ A\, w_i\ =\ \lambda_i(t, t_0)\, w_i
\]
instead.

In \texttt{hadron} this task is performed as follows on the bootstrap
correlator matrix in the most simple case

```{r, cache=TRUE}
t0 <- 4
correlatormatrix.gevp <- bootstrap.gevp(cf=correlatormatrix, t0=t0,
                                        element.order=c(1,2,3,4),
                                        sort.type="values")
```

Next, the principal correlators $\lambda_i$ are obtained as follows,
where in this case we have $i=1,2$

```{r, warning=FALSE}
pc1 <- gevp2cf(gevp=correlatormatrix.gevp, id=1)
pc2 <- gevp2cf(gevp=correlatormatrix.gevp, id=2)
plot(pc1, col="red", pch=21, log="y", xlab="t", ylab="C(t)")
plot(pc2, rep=TRUE, col="blue", pch=22)
```

These principal correlators can be analysed as every object of type
\texttt{cf}, see below. 

## Additional Options

\texttt{bootstrap.gevp} has some additional options which are worth
mentioning.

1. During the bootstrap procedure for the GEVP, eigenvalues have to be
   sorted for every $t$-value. This can be either done by
   \texttt{values}, \texttt{vectors} or \texttt{det} passed via the
   parameter \texttt{sort.type}. When \texttt{vectors} is chosen,
   scalar products of eigenvectors are computed
   \[
   v(t', t_0) \cdot v(t, t_0)
   \]
   and the overlap maximised. When \texttt{sort.t0} is set to
   \texttt{TRUE}, the comparison time is chosen constant as
   $t'=t_0+1$. Otherwise, $t'=t-1$ is set in dependence of $t$. 

2. With parameter \texttt{element.order} the correlation functions in
   the input correlator matrix are specified for use in the GEVP. This
   can be a sub-set of all the correlation functions in the
   matrix. Double usage is allowed as well.
   

# Extracting Energies

```{r}
pc1.effectivemass <- fit.effectivemass(cf=bootstrap.effectivemass(cf=pc1),
                                       t1=5, t2=20)
plot(pc1.effectivemass, col="red", pch=21, ylim=c(0,1))
```

