---
title: "Hankel method for energy level extraction"
author: "Carsten Urbach"
output:
  pdf_document:
    #citation_package: biblatex

#bibliography: hankel.bib
#link_citation: yes
vignette: >
  %\VignetteIndexEntry{Hankel method for energy level extraction}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE}
library(hadron)
```


# Hankel Matrix for a Correlation Function

A $n\times n$ Hankel matrix $H$ corresponding to a vector
$x=(a_0, a_1, a_2, \ldots, a_{2n-2})$ is given by
\begin{equation}
H[x, n] =
\begin{pmatrix}
a_0 & a_1 & a_2 & \ldots & a_{n-1} \\
a_1 & a_2 & a_3 & \ldots & a_{n} \\
a_2 &     &     &        & \vdots \\
\vdots &  &     &        & \vdots \\
a_{n-1} & \ldots &   &   & a_{2n-2} \\
\end{pmatrix}
\end{equation}
Lets for simplicity consider now a single correlation function $C(t)$
for $t = 0, \ldots, T/2$ with $T$ the temporal extent of the
lattice. Define a time shift $\delta t > 0$, an initial time
$t_0\geq 0$ and chose $n < (T/2 - t_0 -\delta t)/2$. Now define two
vectors 
\begin{equation}
\begin{split}
x_1 &= (C(t_0), C(t_0+1), \ldots, C(T/2))\\
x_2 &= (C(t_0+\delta t), C(t_0+\delta t+1), \ldots, C(T/2))\\
\end{split}
\end{equation}
and the following two $n\times n$ Hankel matrices
\begin{equation}
H_1 = H[x_1, n]\,,\qquad H_2 = H[x_2, n]\,.
\end{equation}
With these we can define the following generalised eigenvalue problem
(GEVP)
\begin{equation}
H_2\, v(t_0, \delta t)\ =\ H_1\, \lambda(t_0, \delta t)\, v(t_0, \delta t)
\end{equation}
with eigenvectors $v(t_0, \delta t)$ and eigenvalues $\lambda(t_0, \delta t)$.

If the correlator $C(t)$ is given by a sum of exponentials, i.e.
\[
C(t)\ =\ \sum_{i=1}^n a_i \exp(-E_i t)\,,
\]
one can see that the eigenvalues $\lambda(t_0, \delta t)$ correspond
to the exponentials as follows
\begin{equation}
\lambda_i(t_0, \delta t)\ =\ \exp(-E_i \delta t)\,.
\end{equation}
This method is known as the method of Prony \cite{prony:1795} in the
literature, see also \cite{Lin:2007iq}. For an improved method see \cite{gardner:1959}.

This method is implemented in hadron as follows: we first load the
sample correlator matrix, solve the $4\times 4$ GEVP and determine the first
principal correlator:

```{r}
data(correlatormatrix)
correlatormatrix <- bootstrap.cf(correlatormatrix, boot.R=400, boot.l=1, seed=132435)
correlatormatrix.gevp <- bootstrap.gevp(cf=correlatormatrix, t0=4, element.order=c(1,2,3,4))
pc1 <- gevp2cf(gevp=correlatormatrix.gevp, id=1)
```

Note that the data is for the pion.
Next, we use the first principal correlator as input to the hankel
method. For this, we call

```{r}
pc1.hankel <- bootstrap.hankel(cf=pc1, t0=2, n=2)
```

Thus, $n=2$ and $t_0=2$ in this case. Next, we extract the lowest
eigenvalue by converting into a \texttt{cf} object

```{r}
hpc1 <- hankel2cf(hankel=pc1.hankel, id=1)
```

which looks as follows

```{r}
plot(hpc1, log="y", ylab="lambda(delta t)", xlab="delta t + t0")
```

which could, for instance be analysed using the \texttt{matrixfit}
hadron function. However, we can also cast the eigenvalues directly
into effective masses, since the eigenvalues are generalisations of
those.

```{r}
heffectivemass1 <- hadron:::hankel2effectivemass(hankel=pc1.hankel, id=1)
```

For comparison, we also compute the effective masses of the original
principal correlator

```{r}
pc1.effectivemass <- bootstrap.effectivemass(cf=pc1)
```

and compare in a plot

```{r}
plot(pc1.effectivemass, pch=21, col="red", ylim=c(0,1.1), xlim=c(0,18),
     xlab="t", ylab="M(t)")
plot(heffectivemass1, rep=TRUE, pch=22, col="blue")
legend("topright", legend=c("pc1", "hankel1"), bty="n", pch=c(21,22), col=c("red", "blue"))
```

The result depends strongly on the choice of $t_0$, of course.

```{r, echo=FALSE}
plot(pc1.effectivemass, pch=21, col="red", ylim=c(0,1.1), xlim=c(0,18),
     xlab="t", ylab="M(t)")
plot(heffectivemass1, pch=22, col="blue", rep=TRUE)
pc1.hankel <- bootstrap.hankel(cf=pc1, t0=1, n=2)
heffectivemass1 <- hadron:::hankel2effectivemass(hankel=pc1.hankel, id=1)
plot(heffectivemass1, pch=23, col="darkgreen", rep="TRUE")
pc1.hankel <- bootstrap.hankel(cf=pc1, t0=3, n=2)
heffectivemass1 <- hadron:::hankel2effectivemass(hankel=pc1.hankel, id=1)
plot(heffectivemass1, pch=24, col="black", rep="TRUE")
legend("topright", legend=c("pc1", "t0=2", "t0=1", "t0=3"), bty="n", pch=c(21:24),
       col=c("red", "blue", "darkgreen", "black"))
```

One observes increasing statistical errors, but also earlier plateaus
with increasing $t_0$-values. We can aso apply this method to the
original correlation functions directly without using the GEVP before

```{r}
ppcor <- extractSingleCor.cf(cf=correlatormatrix, id=1)
ppcor.effectivemass <- bootstrap.effectivemass(cf=ppcor)
ppcor.hankel <- bootstrap.hankel(cf=ppcor, t0=3, n=2)
heffectivemass1 <- hadron:::hankel2effectivemass(hankel=ppcor.hankel, id=1)
plot(ppcor.effectivemass, pch=21, col="red", ylim=c(0,1.1), xlim=c(0,18),
     xlab="t", ylab="M(t)")
plot(heffectivemass1, pch=22, col="blue", ylim=c(0,1.1), rep=TRUE)
legend("topright", legend=c("ppcor", "hankel1"), bty="n", pch=c(21,22), col=c("red", "blue"))
```

which works not as well as the method applied to the principal
correlator. 

# Proof

In order to proof the GEVP relation from above, we first introduce a
more general Hankel matrix
\[
H =
\begin{pmatrix}
  s_1 & s_2 & \ldots & s_n\\
  s_2 & s_3 & \ldots &    \\
  \vdots &  &        &    \\
  s_k & \ldots & & s_m\\
\end{pmatrix}
\]
with the signal vector
\[
s_k\ =\ \sum_{i=1}^r c_i z_i^{k-1}\,;\qquad z_j\ =\ e^{\mathrm{i} \omega_j}\,,
\]
(real frequencies $\omega_j$ and coefficients $c_i$) and
\[
k=m-n+1 >  n \geq 1\,.
\]
Define further
\[
e\ =\
\begin{pmatrix}
  1\\
  1\\
  \vdots\\
  1\\
\end{pmatrix}
\]
and
\[
D_c = \mathrm{diag}(c_1, \ldots, c_r)\,,\quad D_z\ =\ \mathrm{diag}(z_1,
\ldots, z_r)\,.
\]
Then, by multiplying out, one can see that
\[
H =
\begin{pmatrix}
  e^T\\
  e^T D_z\\
  \vdots\\
  e^T D_z^{k-1}\\
\end{pmatrix}\cdot
D_c\cdot
(e\  D_z e\ \ldots\ D_z^{n-1} e)\,.
\]
With this one has shown implicitly that the rank of $H$ is $r$.
Now write
\[
H\ =\
\begin{pmatrix}
  g_1 \\
  H_1\\
\end{pmatrix}\ =\
\begin{pmatrix}
  H_2\\
  g_2\\
\end{pmatrix}
\]
with
\[
g_1\ =\
\begin{pmatrix}
  s_1 & s_2 & \ldots & s_n\\
  \vdots &  &        & \vdots \\
  s_{\delta t+1} & & & s_{\delta t+n}\\
\end{pmatrix}
\,,\qquad
g_2\ =\
\begin{pmatrix}
  s_{k-\delta t} &  & \ldots & s_{m-\delta t}\\
  \vdots &  &        & \vdots \\
  s_{k} & & & s_{m}\\
\end{pmatrix}\,.
\]
Define two Vandermonde matrices, which have full rank
\[
V_1\ =\
\begin{pmatrix}
  e^T\\
  e^T D_z\\
  \vdots\\
  e^T D_z^{k-1-\delta t}
\end{pmatrix}\,,\qquad
V_2\ =\
\begin{pmatrix}
  e^T\\
  e^T D_z\\
  \vdots\\
  e^T D_z^{n-1}
\end{pmatrix}\,.
\]
With these we can re-write $H$ as
\[
H\ =\
\begin{pmatrix}
  V_1 \\
  e^T D_z^{k-\delta t}\\
  \vdots\\
  e^T D_z^{k-1}\\
\end{pmatrix}
\cdot D_c\cdot
V_2^T
\]
From this follows
\begin{equation}
H_2\ =\ V_1 D_c V_2^T\,,\quad H_1\ =\ V_1 D_z^{\delta t} D_c V_2^T\,.
\end{equation}
Now, perform a $QR$ decomposition, with $Q$ unitary and $R$ upper
triagular of the Vandermonde matrices $V_i =Q_i R_i\,,\ i=1,2$, which
is always possible due to the full rank property. This 
means 
\[
H_2\ =\ Q_1 R_1 D_c R_2^T Q_2^T\,,\quad H_1\ =\ Q_1 R_1 D_z^{\delta t} D_c R_2^T Q_2^T
\]
and, since $D_z$ is diagonal, by multiplying both equations with
$(R_2^T Q_2^T)^{-1}$ from the right one obtains
\begin{equation}
Q_1 R_1 D_c\ =\ H_2 Q_2 R_2^{-T}\ =\ D_z^{-\delta t} H_1 Q_2 R_2^{-T} \,.
\end{equation}
The last equation is the desired generalised eigenvalue relation with
eigenvalues the diagonal elements of $D_z^{\delta t}$ and the eigenvectors the
columns of the matrix $Q_2 R_2^{-T}$.
