---
title: "Hankel method for energy level extraction"
author: "Carsten Urbach"
output: pdf_document

vignette: >
  %\VignetteIndexEntry{Jackknife Error Normalization}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Hankel Matrix for a Correlation Function

A $n\times n$ Hankel matrix $H$ corresponding to a vector
$x=(a_0, a_1, a_2, \ldots, a_{2n-2})$ is given by
\begin{equation}
H[x, n] =
\begin{pmatrix}
a_0 & a_1 & a_2 & \ldots & a_{n-1} \\
a_1 & a_2 & a_3 & \ldots & a_{n} \\
a_2 &     &     &        & \vdots \\
\vdots &  &     &        & \vdots \\
a_{n-1} & \ldots &   &   & a_{2n-2} \\
\end{pmatrix}
\end{equation}
Lets for simplicity consider now a single correlation function $C(t)$
for $t = 0, \ldots, T/2$ with $T$ the temporal extent of the
lattice. Define a time shift $\delta t > 0$, an initial time
$t_0\geq 0$ and chose $n < (T/2 - t_0 -\delta t)/2$. Now define two
vectors 
\begin{equation}
\begin{split}
x_1 &= (C(t_0), C(t_0+1), \ldots, C(T/2))\\
x_2 &= (C(t_0+\delta t), C(t_0+\delta t+1), \ldots, C(T/2))\\
\end{split}
\end{equation}
and the following two $n\times n$ Hankel matrices
\begin{equation}
H_1 = H[x_1, n]\,,\qquad H_2 = H[x_2, n]\,.
\end{equation}
With these we can define the following generalised eigenvalue problem
(GEVP)
\begin{equation}
H_2\, v\ =\ H_1\, \lambda(t_0, \delta t)\, v
\end{equation}
with eigenvectors $v$ and eigenvalues $\lambda(t_0, \delta t)$.

If the correlator $C(t)$ is given by a sum of exponentials, i.e.
\[
C(t)\ =\ \sum_{i=1}^n a_i \exp(-E_i t)\,,
\]
one can see that the eigenvalues $\lambda(t_0, \delta t)$ correspond
to the exponentials as follows
\begin{equation}
\lambda_i(t_0, \delta t)\ =\ \exp(-E_i \delta t)\,.
\end{equation}
This method is known as the method of Prony \cite{} in the literature.

This method is implemented in hadron as follows: we first load the
sample correlator matrix, solve the $4\times 4$ GEVP and determine the first
principal correlator:

```{r}
data(correlatormatrix)
correlatormatrix <- bootstrap.cf(correlatormatrix, boot.R=400, boot.l=1, seed=132435)
correlatormatrix.gevp <- bootstrap.gevp(cf=correlatormatrix, t0=4, element.order=c(1,2,3,4))
pc1 <- gevp2cf(gevp=correlatormatrix.gevp, id=1)
```

Note that the data is for the pion.
Next, we use the first principal correlator as input to the hankel
method. For this, we call

```{r}
pc1.hankel <- bootstrap.hankel(cf=pc1, t0=2, n=2)
```

Thus, $n=2$ and $t_0=2$ in this case. Next, we extract the lowest
eigenvalue by converting into a \texttt{cf} object

```{r}
hpc1 <- hankel2cf(hankel=pc1.hankel, id=1)
```

which looks as follows

```{r}
plot(hpc1, log="y", ylab="lambda(delta t)", xlab="delta t + t0")
```

which could, for instance be analysed using the \texttt{matrixfit}
hadron function. However, we can also cast the eigenvalues directly
into effective masses, since the eigenvalues are generalisations of
those.

```{r}
heffectivemass1 <- hadron:::hankel2effectivemass(hankel=pc1.hankel, id=1)
```

For comparison, we also compute the effective masses of the original
principal correlator

```{r}
pc1.effectivemass <- bootstrap.effectivemass(cf=pc1)
```

and compare in a plot

```{r}
plot(pc1.effectivemass, pch=21, col="red", ylim=c(0,1.1), xlim=c(0,18))
plot(heffectivemass1, rep=TRUE, pch=22, col="blue")
legend("topright", legend=c("pc1", "hankel1"), bty="n", pch=c(21,22), col=c("red", "blue"))
```
