---
title: "Multi-Particle Fit Test"
author: "Martin Ueding"
date: \today
output:
  pdf_document:
    fig_crop: false
vignette: >
  %\VignetteIndexEntry{Multi-Particle Fit Test}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

devtools::load_all()
```

First we need to create some artificial data in order to get a feeling whether this function actually works. We will take two states and add some noise.

```{r}
time_extent <- 96
e0_val <- 0.12
e1_val <- 0.0
rel_amplitude <- 0.4
boot_R <- 300
n_meas <- 100
```

Next we generate bootstrap samples for the energies and amplitudes that enter here.

```{r}
a0_val <- 1.0
a1_val <- rel_amplitude
a0_err <- 0.001
a1_err <- 0.001
a0_orig <- rnorm(n_meas, a0_val, a0_err)
a1_orig <- rnorm(n_meas, a1_val, a1_err)

e0_err <- 0.001
e1_err <- 0.001
e0_orig <- rnorm(n_meas, e0_val, e0_err)
e1_orig <- rnorm(n_meas, e1_val, e1_err)
```

The signal is just the sum of two cosh-terms.

```{r}
t <- 0:(time_extent - 1)

signal_val <- a0_val * (exp(-e0_val * t) + exp(-e0_val * (time_extent - t))) +
  a1_val * (exp(-e1_val * t) + exp(-e1_val * (time_extent - t)))

signal_orig <- do.call(cbind, lapply(t, function (t) a0_orig * (exp(-e0_orig * t) + exp(-e0_orig * (time_extent - t))) +
  a1_orig * (exp(-e1_orig * t) + exp(-e1_orig * (time_extent - t)))))

signal_orig <- signal_orig + matrix(rnorm(length(signal_orig), c(signal_orig), 0.1), ncol = ncol(signal_orig))

signal_err <- apply(signal_orig, 2, sd)
```

We take a look at our artificial data.

```{r}
plotwitherror(t, signal_val, signal_err,
     log = 'y',
     main = 'Artificial pure signal')
```

All we have are the fake measurements, so we construct a `cf` object from them and bootstrap that.

```{r}
corr <- cf_orig(cf_meta(Time = time_extent), cf = signal_orig)
corr_boot <- bootstrap.cf(symmetrise.cf(corr), boot.R = boot_R)
```

Now it looks like a correlator with very tiny errors.

```{r}
plot(corr_boot,
     log = 'y',
     main = 'Symmetrized correlator with noise')
```

In the effective mass we can see that we are hopelessly far away from the actual mass that I want to get out.

```{r}
effmass_solve <- bootstrap.effectivemass(corr_boot, type = 'solve')
plot(effmass_solve, ylim = range(effmass_solve$t0, e0_val, na.rm = TRUE))
abline(h = e0_val)
```

Shifting the correlator gives us a sensible effective mass because it removes the constant term.

```{r}
corr_shifted <- takeTimeDiff.cf(corr_boot)
effmass_shifted <- bootstrap.effectivemass(corr_shifted, type = 'shifted')
plot(effmass_shifted)
abline(h = e0_val)
```

# Fits

We first try the “single” model, that will of course fail miserably.

```{r}
fit_single <- new_matrixfit(
  corr_boot,
  t1 = 1,
  t2 = 47,
  useCov = TRUE,
  model = 'single',
  fit.method = 'lm')

plot(fit_single,
     log = 'y',
     main = 'Fit with “single” model')
residual_plot(fit_single)
```

The “shifted” model will of course work.

```{r}
fit_shifted <- new_matrixfit(
  corr_shifted,
  t1 = 1,
  t2 = 47,
  model = 'shifted',
  useCov = TRUE,
  fit.method = 'lm')

plot(fit_shifted,
     log = 'y',
     main = 'Fit with “shifted” model')
residual_plot(fit_shifted)
```

And with the “n_particle” model we can also get the result. It is not entirely correct because I must not use the original data for $E_1$ but rather the bootstrapped one. But I do not want to bootstrap that vector since I lack a function to do it nicely for me. So I just generate it uncorrelated.

```{r}
fit_n_particle <- new_matrixfit(
  corr_boot,
  t1 = 1,
  t2 = 47,
  model = 'n_particles',
  fit.method = 'lm',
  useCov = TRUE,
  higher_states = list(val = c(e1_val),
                       ampl = c(1),
                       boot = matrix(rnorm(boot_R, e1_val, e1_err / sqrt(n_meas)), ncol = 1)))

plot(fit_n_particle,
     log = 'y',
     main = 'Fit with “n_particle” model')
residual_plot(fit_n_particle)
```
