---
title: "PGEVM and Lanczos with double bootstrap"
author: "Carsten Urbach"
output:
  rmarkdown::html_vignette
  
#  pdf_document:
#    #citation_package: biblatex

#bibliography: hankel.bib
#link_citation: yes
vignette: >
  %\VignetteIndexEntry{PGEVM method for energy level extraction}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE}
library(hadron)
```

# The Standard Effective Mass analysis

'hadron' ships with a pion correlation function, which we load now
first. Then we store it in an object of class 'cf' (the hadron
correlation function class) for further treatment.

```{r}
data(pscor.sample)
pioncf <- cf_orig(cf=t(array(pscor.sample[,2], dim=c(48, 316))))
pioncf <- cf_meta(pioncf, nrObs=1, Time=48, symmetrised=FALSE)
```

Next we (block) bootstrap and double bootstrap this correlation
functions. Note that depending on the choice of 'boot.R' and
'dbboot.R' the execution of the following lines might take long. For a
fast test set both of them to for instance '20'.

```{r}
boot.R <- 29
dbboot.R <- 19
pioncf.boot <- bootstrap.cf(pioncf, boot.R=boot.R, boot.l=2)
pioncf.boot <- double_bootstrap.cf(pioncf.boot, dbboot.R=dbboot.R)
```

For the standard effective mass 

```{r}
pion.efm <- bootstrap.effectivemass(pioncf.boot, type="solve")
pion.efm <- fit.effectivemass(pion.efm, t1=8, t2=17, useCov=TRUE)
```

which looks as follows

```{r}
plot(pion.efm, ylab="Meff", xlab="t/a", xlim=c(0,24), ylim=c(0.1,0.2))
```

# PGEVM

Based on the (double) bootstrap correlation function data, the PGEVM
is called as follows for double bootstrap:

```{r}
pion.pgevm <- bootstrap.pgevm(pioncf.boot, N=pioncf.boot$Time-2)
pion.pgevm.efm <- pgevm2effectivemass(pion.pgevm, errortype="dbboot", bias_correction=TRUE)
```

Plotting this works in an identical manner as the effective mass before

```{r}
plot(pion.pgevm.efm, ylab="Meff", xlab="n", xlim=c(0,22), ylim=c(0.1,0.2), col="blue")
```

Also otherwise 'pion.pgevm.efm' can be treated like a class of type
'effectivemass'. For instance, one can also fit:

```{r}
pion.pgevm.efm <- fit.effectivemass(pion.pgevm.efm, t1=10, t2=21, useCov=TRUE)
```

where now 't1' and 't2' are interpreted as $n_1$ and $n_2$. The
corresponding plot looks as follows

```{r}
plot(pion.pgevm.efm, ylab="Meff", xlab="n", xlim=c(0,22), ylim=c(0.1,0.2), col="blue")
```

The fitresult reads $E =`r pion.pgevm.efm$effmassfit$t0[1]` \pm `r pion.pgevm.efm$effmassfit$se[1]`$.

# Lanczos

Lanczos works similarly. It is applied to the double bootstrapped
correlation data with the function 'bootstrap.lanczos' as follows

```{r}
pion.lanczos <- bootstrap.lanczos(pioncf.boot, N=pioncf.boot$Time-2,
                                  bias_correction=TRUE, errortype="dbboot", pivot=TRUE)
```

and can can be plotted similarly

```{r}
plot(pion.lanczos, ylab="Meff", xlab="n", xlim=c(0,22), ylim=c(0.1,0.2), col="red")
```
