---
title: "PGEVM and Lanczos with double bootstrap in hadron"
subtitle: "Companion document to arXiv:2411.14981"
author: 
- Johann Ostmeyer
- Aniket Sen 
- Carsten Urbach
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmarkdown::html_vignette
  
#  pdf_document:
#    #citation_package: biblatex

#bibliography: hankel.bib
#link_citation: yes
vignette: >
  %\VignetteIndexEntry{PGEVM and Lanczos with double bootstrap in hadron}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

This is a companion document for the manuscript
[arxiv.org/abs/2411.14981](https://arxiv.org/abs/2411.14981). It
explains the PGEVM and Lanczos implementations available in the
`hadron` `R` package. The `hadron` package version required is
[version 3.3.1](https://github.com/HISKP-LQCD/hadron/tree/3.3.1) or newer.
It is available on `github` for download. 

You first need to install the `hadron` package, and load it then as follows

```{r}
library(hadron)
```

The code snippets given below can be copied one-to-one to the `R`
command line and executed. Alternatively, you may install the
`rmarkdown` package and render the corresponding `Rmd` file, which was
used to generate this PDF file as follows

```{r, eval=FALSE}
render("pgevm.Rmd", output_format="pdf_document")
```

`hadron` ships with a pion correlation function, which was also used
as the pion example in the aforementioned manuscript.

# The Standard Effective Mass analysis

After loading `hadron` as a library, one first needs to load the pion
correlator data

```{r}
data(pscor.sample)
```

and store it in an object of class `cf` (the hadron correlation
function class) for further treatment:

```{r}
pioncf <- cf_orig(cf=t(array(pscor.sample[,2], dim=c(48, 316))))
pioncf <- cf_meta(pioncf, nrObs=1, Time=48, symmetrised=FALSE)
```

Next, we (block) bootstrap and double bootstrap this correlation
function. Note that depending on the choice of `boot.R` and
`dbboot.R` the execution of the following lines might take long.

```{r}
boot.R <- 29
dbboot.R <- 19
pioncf.boot <- bootstrap.cf(pioncf, boot.R=boot.R, boot.l=2)
pioncf.boot <- double_bootstrap.cf(pioncf.boot, dbboot.R=dbboot.R)
```

`hadron` provides functionality to compute the standard effective mass
for a $\\cosh$-like correlation function (specified by the `type`
argument to the function `bootstrap.effectivemass`)

```{r}
pion.efm <- bootstrap.effectivemass(pioncf.boot, type="solve")
pion.efm <- fit.effectivemass(pion.efm, t1=8, t2=17, useCov=TRUE)
```

Note that you may get documentation on the different functions e.g. by

```{r, eval=FALSE}
?bootstrap.effectivemass
```

The such computed effective masses and the corresponding fit can be
visualised as follows

```{r}
plot(pion.efm, ylab="Meff", xlab="t/a", xlim=c(0,24), ylim=c(0.1,0.2))
```

A call to 

```{r, eval=FALSE}
summary(pion.efm)
```

would also summarize the fit results.

# PGEVM

Based on the (double) bootstrapped correlation function data, the PGEVM
is applied as follows:

```{r, warn=FALSE}
pion.pgevm <- bootstrap.pgevm(pioncf.boot, N=pioncf.boot$Time-1)
pion.pgevm.efm <- pgevm2effectivemass(pion.pgevm, errortype="dbboot",
                                      bias_correction=TRUE)
```

Note that you might get one or more warning from the above commands,
which are due to the fact that for some combination of data the Hankel
matrix cannot be inverted or the eigenvalue compuation fails. But this
is taken into account in the analysis by missing values.

The second line translates the PGEVM result into an effective mass
object. Thus, plotting it works in an identical manner as for the true
effective mass

```{r}
plot(pion.pgevm.efm, ylab="Meff", xlab="n", xlim=c(0,22), ylim=c(0.1,0.2), col="blue")
```

Also otherwise, `pion.pgevm.efm` can be treated like a class of type
`effectivemass`. For instance, one can also perform a plateau fit:

```{r}
pion.pgevm.efm <- fit.effectivemass(pion.pgevm.efm, t1=10, t2=21, useCov=TRUE)
```

where now `t1` and `t2` are interpreted as $n_1$ and $n_2$. The
corresponding plot looks as follows

```{r}
plot(pion.pgevm.efm, ylab="Meff", xlab="n", xlim=c(0,22), ylim=c(0.1,0.2), col="blue")
```

The fit result reads $E =`r pion.pgevm.efm$effmassfit$t0[1]` \pm `r pion.pgevm.efm$effmassfit$se[1]`$.

# Lanczos

Lanczos works similarly. It is applied to the double bootstrapped
correlation data with the function `bootstrap.lanczos` as follows

```{r}
pion.lanczos <- bootstrap.lanczos(pioncf.boot, N=pioncf.boot$Time-1,
                                  bias_correction=TRUE, errortype="dbboot", pivot=TRUE)
```

and can can be plotted similarly

```{r}
plot(pion.lanczos, ylab="Meff", xlab="n", xlim=c(0,22), ylim=c(0.1,0.2), col="red")
```

# PGEVM for periodic correlators

```{r, warn=FALSE}
pion.pgevm2 <- bootstrap.pgevm(pioncf.boot, N=pioncf.boot$Time-1, ndep.Delta=TRUE)
pion.pgevm.efm2 <- pgevm2effectivemass(pion.pgevm2, errortype="dbboot",
                                       bias_correction=TRUE)
pion.pgevm.efm3 <- pgevm2effectivemass(pion.pgevm2, errortype="dbboot",
                                       bias_correction=TRUE, average.negE=TRUE)
```

```{r}
plot(pion.pgevm.efm, ylab="Meff", xlab="n", xlim=c(0,22), ylim=c(0.1,0.2), col="blue", pch=21)
plot(pion.pgevm.efm2, rep=TRUE, col="red", pch=22, xshift=0.2)
plot(pion.pgevm.efm3, rep=TRUE, col="darkorange", pch=23, xshift=0.4)
```
