---
title: "Truncated PGEVM in hadron"
subtitle: "Companion document to arXiv:2509.xxxxx"
author: 
- Johann Ostmeyer
- Carsten Urbach
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmarkdown::html_vignette
  
#  pdf_document:
#    #citation_package: biblatex

#bibliography: hankel.bib
#link_citation: yes
vignette: >
  %\VignetteIndexEntry{Truncated PGEVM in hadron}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

This is a companion document for the manuscripts
[arxiv.org/abs/2411.14981](https://arxiv.org/abs/2411.14981) and work forthcoming. It
explains the truncated PGEVM implementation available in the
`hadron` `R` package. The `hadron` package version required is
[version 3.4.1](https://github.com/HISKP-LQCD/hadron/tree/3.4.1) or newer.
It is available on `github` for download. 

You first need to install the `hadron` package, and load it then as follows

```{r}
library(hadron)
```

The code snippets given below can be copied one-to-one to the `R`
command line and executed. Alternatively, you may install the
`rmarkdown` package and render the corresponding `Rmd` file, which was
used to generate this PDF file as follows

```{r, eval=FALSE}
render("truncated_pgevm.Rmd", output_format="pdf_document")
```

`hadron` ships with a pion correlation function, which was also used
as the pion example in the aforementioned manuscript.

# The Standard Effective Mass analysis

After loading `hadron` as a library, one first needs to load the pion
correlator data

```{r}
data(pscor.sample)
```

and store it in an object of class `cf` (the hadron correlation
function class) for further treatment:

```{r}
T <- 48
pioncf <- cf_orig(cf=t(array(pscor.sample[,2], dim=c(T, 316))))
pioncf <- cf_meta(pioncf, nrObs=1, Time=T, symmetrised=FALSE)
```

Next, we (block) bootstrap and double bootstrap this correlation
function. Note that depending on the choice of `boot.R` and
`dbboot.R` the execution of the following lines might take long.

```{r}
boot.R <- 29
dbboot.R <- 19
pioncf.boot <- bootstrap.cf(pioncf, boot.R=boot.R, boot.l=2)
pioncf.boot <- double_bootstrap.cf(pioncf.boot, dbboot.R=dbboot.R)
plot(pioncf.boot, ylab="C(t)", xlab="t/a", xlim=c(0,T), log="y")
```

Here, we have also plotted the bootstrapped correlation function using
the inbuilt `plot` function for objects of class `cf`. The correlator
is symmetric around $T/2=24$ as expected.

`hadron` provides functionality to compute the standard effective mass
for a $\\cosh$-like correlation function (specified by the `type`
argument to the function `bootstrap.effectivemass`)

```{r}
pion.efm <- bootstrap.effectivemass(pioncf.boot, type="solve")
pion.efm <- fit.effectivemass(pion.efm, t1=8, t2=17, useCov=TRUE)
```

Note that you may get documentation on the different functions e.g. by

```{r, eval=FALSE}
?bootstrap.effectivemass
```

The such computed effective masses and the corresponding fit can be
visualised as follows

```{r}
plot(pion.efm, ylab="Meff", xlab="t/a", xlim=c(0,T/2), ylim=c(0.1,0.2))
```

A call to 

```{r, eval=FALSE}
summary(pion.efm)
```

would also summarize the fit results.

# Basic PGEVM

Based on the (double) bootstrapped correlation function data, the PGEVM
is applied as follows:

```{r, warn=FALSE}
pion.pgevm <- bootstrap.pgevm(pioncf.boot, N=pioncf.boot$Time)
pion.pgevm.efm <- pgevm2effectivemass(pion.pgevm, errortype="dbboot",
                                      bias_correction=TRUE)
```

Note that you might get one or more warning from the above commands,
which are due to the fact that for some combination of data the Hankel
matrix cannot be inverted or the eigenvalue compuation fails. But this
is taken into account in the analysis by missing values.

The second line translates the PGEVM result into an effective mass
object. Thus, plotting it works in an identical manner as for the true
effective mass

```{r}
plot(pion.pgevm.efm, ylab="Meff", xlab="n", xlim=c(0,22), ylim=c(0.1,0.2), col="blue")
```

Also otherwise, `pion.pgevm.efm` can be treated like a class of type
`effectivemass`. For instance, one can also perform a plateau fit:

```{r}
pion.pgevm.efm.fit <- fit.effectivemass(pion.pgevm.efm, t1=10, t2=21, useCov=TRUE)
```

where now `t1` and `t2` are interpreted as $n_1$ and $n_2$. The
corresponding plot looks as follows

```{r}
plot(pion.pgevm.efm.fit, ylab="Meff", xlab="n", xlim=c(0,22), ylim=c(0.1,0.2), col="blue")
```

The fit result reads $E =`r pion.pgevm.efm.fit$effmassfit$t0[1]` \pm `r pion.pgevm.efm.fit$effmassfit$se[1]`$.

# PGEVM for symmetric correlators

For symmetric correlators it makes sense to include not only early but also late time slices starting from low dimensions. This can be done by setting `ndep.Delta=TRUE` in the `bootstrap.pgevm` function and thus distributing the correlator time slices entering the Hankel matrix equidistantly. The following code snippet demonstrates this:

```{r, warn=FALSE}
pion.pgevm2 <- bootstrap.pgevm(pioncf.boot, N=pioncf.boot$Time, ndep.Delta=TRUE)
pion.pgevm.efm2 <- pgevm2effectivemass(pion.pgevm2, errortype="dbboot",
                                       bias_correction=TRUE)
pion.pgevm.efm3 <- pgevm2effectivemass(pion.pgevm2, errortype="dbboot",
                                       bias_correction=TRUE, average.negE=TRUE)
```

In the last step the `average.negE=TRUE` option is used to average over the ''effective masses'' that are now symmetric around zero.

We can now plot the different effective masses.

```{r}
plot(pion.pgevm.efm, ylab="Meff", xlab="n", xlim=c(0,T/2), ylim=c(.12,.155),
                     col="blue", pch=21, xshift=0)
plot(pion.pgevm.efm2, rep=TRUE, col="red", pch=22, xshift=0.15)
plot(pion.pgevm.efm3, rep=TRUE, col="darkorange", pch=23, xshift=0.3)
```

# Truncated PGEVM

Instead of going time slice by time slice, we can also include all the
correlator values at once and truncate the diagonalised Hankel matrix. This is
done as follows:

```{r, warn=FALSE}
t0 <- 1
pion.truncated.pgevm <- bootstrap.truncated.pgevm(pioncf.boot, N=pioncf.boot$Time,
                                      t0=t0, n=T/2-t0, error.weights=TRUE, symmetric=TRUE)
pion.truncated.pgevm.efm <- pgevm2effectivemass(pion.truncated.pgevm,
                                      errortype="outlier-removal", bias_correction=TRUE)
```

Here, we set the initial time $t_0 = `r t0`$ as early as possible and the full
Hankel dimension $n = T/2 - t_0 = `r T/2-t0`$ maximal to include all time
slices from $t=t_0$ to $t=t_0+2n=`r T-t0`$, symmetric about $T/2=`r T/2`$. The
`error.weights=TRUE` option ensures that the Hankel matrix eigenspectrum is
computed using the inverse of the variance as weights. The `symmetric=TRUE`
option takes into account that the correlator is symmetric around $T/2$.

Note that the results are stable enough that simple outlier removal is
sufficient to obtain a reliable error estimate. Double bootstrap would also
work, but is more expensive.

It is insightful to plot the absolute values of the eigenvalues of the Hankel
matrix:

```{r}
plot(abs(pion.truncated.pgevm$singular.values), log="y", ylab="Eigenvalues")
points(-Re(pion.truncated.pgevm$singular.values), col="red")
legend("topright", legend=c("positive", "negative"), col=c("black", "red"), pch=1)
trunc <- pion.truncated.pgevm$opt.idx[1]
```

Clearly, there are $`r trunc`$ dominant eigenvalues, which we will keep in the
following. The remaining part of the spectrum stems from noise.
The variable `opt.idx` contains the maximal logarithmic gap in the spectrum
(used here) and the position before the first negative eigenvalue (because the
noiseless Hankel matrix would be positive semi-definite).

The (close-to) optimal fit result after truncation at $k = `r trunc`$ states
reads
$E =`r pion.truncated.pgevm.efm$t0[trunc]` \pm `r pion.truncated.pgevm.efm$se[trunc]`$.

Once again, we can now plot the different "effective masses". It is crucial to
keep in mind that the results from the truncated PGEVM are not independent for
different $n$ as the same Hankel matrix is used everywhere. Thus, a better way
to interpret the results is to identify every truncation $k$ with a multi-state
fit using $k$ states. Any individual point already contains the full
statistical information. The resulting "plateau" allows to verify that there is
no significant systematic bias from the truncation.

```{r}
plot(pion.pgevm.efm, ylab="Meff", xlab="n or k", xlim=c(0,T/2), ylim=c(.12,.155),
                     col="blue", pch=21, xshift=0)
plot(pion.pgevm.efm3, rep=TRUE, col="darkorange", pch=23, xshift=0.15)
plot(pion.truncated.pgevm.efm, rep=TRUE, col="red", pch=22, xshift=0.3)
```

The truncated PGEVM also allows us to reconstruct the coefficients of the
exponential fit ansatz and to compare the result with the original data. This
can be done as follows:

```{r}
pion.pgevm.nlsfit <- pgevm2bootstrapfit(pion.truncated.pgevm, truncation.dim=trunc)
plot(pion.pgevm.nlsfit[[1]], log="y", xlab="t", ylab="C(t)", col.line="red", col.band="red")
residual_plot(pion.pgevm.nlsfit[[1]], ylim=c(.9,1.1), xlab="t", ylab="C(t)/fit")
```
